Seed ecology of European mesic meadows: the domestication of plant regeneration strategies

### 2021-06-30 doc/manuscript.Rmd
pc mig

European mesic meadows are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698]. These plant communities developed in clearings of a temperate woodland landscape as a consequence of anthropogenic intervention [@RN5033; @RN5045]. At a time, they were novel plant communities with a composite flora made up of species from different origins: grasslands plants that existed in non-forested "islands" of the landscape, plus plants that evolved in parallel to human intervention by hybridization or polyploidization [@RN5033]. During historical times the botanical composition of mesic meadows evolved following changes in human activities and management practices [@RN5035; @RN5045]. They are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034]. Nowadays, mesic meadows maintained by agricultural practices are an essential aspect of European landscapes, both real [@RN4996] and imagined [@RN5029].

Traditional phytosociological classification of mesic grasslands puts emphasis on the separation between meadows used for hay making versus pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893]. Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic meadows in large parts of Europe [@RN4997]. For this reason, the European Habitats Directive (92/43/EEC) has included in its lists of habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.

@@

Mesic meadows are dominated by mesophilous grasses of the Poaceae family, which make up most of the biomass and define the structure of the vegetation. The dispersal unit in the grasses is generally the floret (**Fig. 1a**), a composite structure made up of the caryopsis or grain (a dry, indehiscent and monospermic fruit) and its surrounding bracts (i.e. modified leaves): the lemma and the palea. Removal of the lemma and palea, as well as puncturing the pericarp, can aid in seed germination [@RN1248]. The dispersal unit is often accompanied by awns or hairs (**Fig. 1b**) that are assumed to aid dispersal, although it is not always the case that they do [@RN1290]. Meadow grass seeds have been reported as having physiological seed dormancy [@RN5018; @RN3214; @RN5024] in various degrees, but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273]. Freshly harvested seeds are comparatively more dormant, but dormancy tends to disappear quickly in dry storage [@RN5024; @RN5018]. Germination has been reported to occur at temperatures ranging from 5 to 30 ºC [@RN3273; @RN1229; @RN1261; @RN1230; @RN1292; @RN5024]. Most Poaceae species have also been reported to germinate better in light than in darkness [@RN1271; @RN1248; @RN1230; @RN1221; @RN1172; @RN1292; @RN1280; @RN5024] and even to be inhibited to germinate by darkness [@RN1229]. Nonetheless, species of *Bromus* have been reported as germinating better in darkness [@RN1172] and *Cynosurus cristatus* as being indifferent to light/darkness [@RN1292; @RN1280]. In *Poa trivialis*, germination was promoted by light but not by alternating temperatures [@RN1205], although alternating temperatures did encourage some germination in darkness [@RN1229]. Wild grass ecotypes usually have a germination response to alternating temperatures [@RN1248; @RN1261; @RN1230; @RN1221; @RN1290; @RN1172; @RN1292; @RN1280], although this response is missing in some wild ecotypes and in the domesticated cultivars [@RN1261], and some species such as *Lolium perenne* have been reported as insensitive to temperature alternation [@RN1280; @RN1380].

Next in abundance to the grasses are the legumes of the Fabaceae family. Legumes contribute to the nutritional value of meadow fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040]. The dispersal unit of most legumes is the seed itself (**Fig. 1c**) but in some species dispersal units are more complex, including indehiscent monospermic fruits (e.g. *Onobrychis*, **Fig. 1d**) or indehiscent monospermic fruit fragments, i.e. loments (e.g. *Ornithopus*, **Fig. 1e**). Legume seeds are generally hard-seeded, i.e. they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e. physical seed dormancy) [@RN3214; @RN3273; @RN1215; @RN5016; @RN3073]. In *Medicago*, seeds that have not reached full maturity can germinate before they become impermeable, but the completion of maturation imposes coat impermeability, and thereafter the seed must be scarified to allow water imbibition and germination [@RN652]. Buried legume seeds are able to track the seasons, and in some species germination seems to be promoted by cold stratification and alternating temperatures [@RN780]; some of these species have been described as having combinational dormancy (i.e. physical + physiological) [@RN5037]. However, seeds have also been reported to germinate without any previous treatment [@RN881; @RN632; @RN460; @RN372], and to lose dormancy during storage [@RN5037]. As in the grasses, legume seeds have been reported to germinate in high numbers across a range of temperatures from 5 to 25 ºC [@RN652; @RN3273]. Legume seeds have been described as not responsive to light and capable of germinating in darkness [@RN1339; @RN3273].

Grasses and legumes are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the aesthetic value of the meadows [@RN5038; @RN5039]. They also add nutritional scope, being richer than grasses and legumes in specific elements [@RN5028; @RN5040]. In many of the most frequent minor families (e.g. in the Asteraceae, Ranunculaceae, Polygonaceae, Lamiaceae, Dipsacaceae, Cyperaceae) the dispersal unit is the achene: dry, indehiscent and monospermic fruits. The morphology of these achenes is varied: cypselae with a hairy pappus in *Centaurea* (Asteraceae, **Fig. 1f**); beaked in *Ranunculus* (Ranunculaceae, **Fig. 1g**); hardened nutlets in *Prunella* (Lamiaceae, **Fig. 1h**); hairy in *Knautia* (Dipsacaceae, **Fig. 1i**); surrounded by a perigynium which aids in dispersal by water in *Carex* (Cyperaceae, **Fig. 1j**). In the genus *Sanguisorba* (Rosaceae), the dispersal unit is the urn-shaped receptacle containing one to three achenes (**Fig. 1k**). In the Apiaceae, it is the mericarp (**Fig. 1l**), an indehiscent monospermic fragment of the fruit. In some other minor families the dispersal unit is the seed itself, such as in the Caryophyllaceae (**Fig. 1m**), the Plantaginaceae (**Fig. 1n**), the Juncaceae or the hemiparasitic species of *Rhinanthus* (Orobanchaceae) (**Fig. 1o**). In the Asteraceae, high germination without pretreatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; @RN120; @RN963; @RN1428], *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403]. *Taraxacum officinale* germinates between 5 and 30 ºC [@RN120; RN1428; @RN1276], and has higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428]. In *Stachys officinalis* (Lamiaceae), seeds need either cold stratification, light or alternating temperatures to germinate [@RN416; @RN128]. Underdeveloped embryos that need to grow inside the seed before germination (i.e. morphological dormancy) are widespread in the Ranunculaceae and the Apiaceae [@RN3214; @RN1112]. *Ranunculus repens* (Ranunculaceae) germinates between 10 and 25 ºC, but the germination percentages were reported to be low [@RN931]; the same species has been reported to respond to alternating temperatures, which can promote its germination even in darkness [@RN3180]. In the Polygonaceae, *Rumex acetosa* can germinate immediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not, this difference is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814]. In *Heracleum sphondylium* (Apiaceae), growth of the embryo only occurs below 10 ºC, in moist conditions [@RN1112]. *Sanguisorba minor* (Rosaceae) increases its germination after abrasion of the seeds with bleach [@RN194; @RN226], although germination without pre-treatment has also been reported [@RN298; @RN226]. Seeds of the hemiparasitic species *Rhinanthus angustifolius* and *Rhinanthus minor* (Orobanchaceae) require relatively long periods of cold stratification to germinate [@RN4014], and are able to germinate in the dark [@RN2236].

Although a wealth of studies has accumulated, a synthesis of the seed ecology of mesic meadows is missing, despite their conservation, restoration and cultural significance. In this article, we present for the first time a synthesis of regeneration by seed in European mesic meadows. To do so, we combine newly-generated data on seed morphology and germination with records from databases [@RN5036; @RN4968; @RN2987; @RN2319]. To the resulting dataset containing `r MSrecords` germination records from `r MScountries` European countries we apply Bayesian meta-analysis [@RN5000] to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light. We analyze the relationship between seed traits and species environmental preferences, and compare the germination ecology of mesic meadows with that of other herbaceous plant communities.

@@

Sampled meadows had been maintained by traditional agricultural practices: mowing for hay-making once or twice per year plus light grazing. Each sampling plot was placed in a square area (25-100 m^2^ area) situated in the central part of a meadow, avoiding the margins. Sampling took place in 2016-2017, at the peak of plant development, just before mowing. All vascular plant species in the plots were recorded and given a cover value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%). All plant names were assigned following the nomenclature of Euro+Med [-@RN5044], which is used throughout this article. As expected, the vegetation of the sampled meadows was dominated by the Poaceae and the Fabaceae: these two families represented 47% and 17%, respectively, of the total plant cover recorded in all the plots. Other 20 families were recorded, the largest of which was the Asteraceae; each of these other families represented less than 10% of the total cover. These numbers refer only to the seed plants, pteridophytes being a marginal part of the records.

Using all the relevés, we calculated the cumulative cover of each species in the entire area. To perform the calculation, first we standardized the cover values of the plots by dividing the cover of each species in each plot by the total plant cover in that plot. Then, for each species, we calculated its total cover in the dataset, by summing its standardized cover values from all the plots. Finally, we rescaled the values of all species to a 1-100 scale to obtain the cumulative cover values. We removed 208 species with cumulative cover values below 2%, considering them to be transient species that might have been recorded by chance and may not represent the core mesic meadows flora. We used the remaining 116 species as the core list to retrieve seed germination data for this article. The majority of these species were hemicryptophytes (78%), with some therophytes (16%) and a few chamaephytes and geophytes (3% of each). The family with more species in the list was the Poaceae (22%), with another 17% belonging to the Fabaceae, 15% to the Asteraceae, 8% to the Apiaceae, and the rest of the families representing less than 5% each.

We also used the relevés as a basis to characterize the preferences of the selected species for three ecological drivers of mesic meadow diversity: cold, summer drought and soil reaction [@RN2980]. For cold and drought, we used the coordinates of the plots to retrieve from CHELSA [@RN4766] the bioclimatic variables bio06 (minimum temperature of the coldest month) and bio14 (precipitation of the driest month). and represents the limiting water availability along the Iberian gradient of summer drought. For soil reaction, we took from each plot five soil samples from between 0 and 20 cm depth with a Dutch auger and combined them to make a bulk soil horizon, which we subsequently air-dried, crumbled, finely crushed and sieved with a 2 mm screen, to finally measure the pH in H~2~O with a glass electrode in a suspension of soil and water (1:2.5). For each of these three environmental variables (bio06, bio14, pH) measured at the plot level we calculated the species niche centroids (SNCs). The SNC for any given species and variable is the mean of the environmental variable in all the plots were the species occurs, weighted by species cover in each plot [@RN4709]. The list of core species, with their cumulative covers and SNCs, is available at GitHub (see Data Availability Statement).

@@

We germinated the collected seeds using three germination treatments to determine the germination response to temperatures that are representative of the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as the summer temperature, and 30/20 ºC as sun-heated soil, e.g. soil exposed to sun after hay cutting. Additionally, we compared, at each of these temperature regimes, the germination of fresh seeds versus seeds subjected to a dormancy-breaking treatment. In the case of the Fabaceae and other families that might present physical dormancy [@RN3214], the treatment consisted in scarification by chipping the seed coat with a scalpel. For the rest of the families, we used gibberellic acid GA~3~ (0,0645 mM) in darkness during 24h, as a treatment to remove potential physiological seed dormancy [@RN4976]. For each species and treatment, we sowed four Petri dishes with 25 seeds each. The germination substrate was 1% distilled water - agar. We sealed dishes with Parafilm to prevent desiccation. Trials took place in a germination chamber (KBW 400, Binder GmbH, Tuttlingen, Germany) with a 12/12-hour photoperiod (the light period corresponding to the higher temperature). Experiments lasted for four weeks, with germination scoring once per week. The germination criterion was 2 mm radicle emergence. After four weeks, we cut the seeds that failed to germinate and examined them under a magnifying glass. We classified them as normal when the embryo was visible and firm, empty when they lacked an embryo, and contaminated when they were moldy. We only considered normal seeds to calculate germination proportions and conduct subsequent analyses.

@@

The Poaceae had lower values of seed mass and higher values of seed number, while the Fabaceae had heavier but fewer seeds (**Fig. 2**). The other families covered the range of values showed by the Poaceae and the Fabaceae, but their median values were high for both traits: their median seed mass was close to that of the Fabaceae, while their median seed number was higher than that of the Poaceae (**Fig. 2**). Seed shape also showed a divergence between the Poaceae and Fabaceae, with seeds being elongated in the former and rounder in the latter (**Fig. 2**). The other families covered the full range of variation, with both elongated and round seeds (**Fig. 2**).

@@

Some differences became apparent when dividing the dataset in the three floristic groups that compose mesic meadow vegetation. In the dominant group of the Poaceae (**Fig. 3**), no effect was found for stratification, GA~3~ or temperature. Stratification had a negative effect on germination. The major drivers of Poaceae germination appeared to be alternating temperatures and light, with both having a positive effect. In the group of the Fabaceae (**Fig. 3**), the largest positive effect on germination was produced by scarification, with no effect of stratification, GA~3~ or light. Temperature had a negative effect, and in fact Fabaceae species had higher germination proportion at temperatures under 20 ºC. Fabaceae seed germination also responded negatively to alternating temperatures. Finally, in the remaining families (**Fig. 3**), the main cues having a positive effect on germination were stratification, GA~3~, alternating temperatures and light. These species did not respond to scarification nor to average temperature.

@@

PCA indicated a clear separation between environmental preferences and seed traits, with each set of variables contributing to different axes (**Fig. 4**). The first PCA axis explained `r round(pca$eig[1, 2], 0)`% of the variation and was related to environmental preferences. The variables with the largest contribution to this first axis were soil reaction (pH), winter cold (bio06) and summer drought (bio14). This horizontal axis separated (i, left) species with preferences for sites with warm winter temperatures from (ii, right) species with preferences for sites with high summer rainfall and alkaline soils. Axis 2 explained `r round(pca$eig[2, 2], 0)`% of the variability and related to seed traits. The main contributing variables were seed mass, seed number and the germination response to scarification. This axis separated (iii, bottom) species of the Poaceae that produced many seeds with a positive germination response to alternating temperatures from (iv, top) species of the Fabaceae that produced heavy seeds with a positive response to scarification.

@@

The second ordination resulted in a separation of mesic meadow species from species belonging to natural herbaceous communities of the same region (**Fig. 5**). Axis 1 explained 78% of the variance and was positively associated to germination at all three temperature treatments. This horizontal axis separated (i, left) species with low germination across treatments from (ii, right) species with high germinability across treatments. The second axis explained 17% of the variation and separated (iii, bottom) species that responded more to the cool temperature germination treatment from (iv, top) species that responded more positively to the warm germination treatment. Meadow species tended to be situated to the right in the horizontal axis, indicating high germinability; and at the center of the vertical axis, indicating a neutral response to temperature. On the other hand, species from the other communities were positioned at the left of the horizontal axis, indicating a lower germinability; and were more separated along the vertical axis, indicating a preference for either warmer (bogs and fens) or cooler (alpine grasslands, coastal communities) germination treatments.

@@

Our meta-analysis of germination records showed that, overall, the regeneration of mesic meadows is characterized by (1) high germinability when compared to other herbaceous plant communities; (2) low variation in relation to natural environmental drivers of mesic meadow diversity; and (3) a strong phylogenetic separation between the two major families, the Poaceae and the Fabaceae.

Alternating temperatures and light are usually considered to be micro-environmental cues that indicate, at a fine scale, the existence of safe sites for regeneration [@RN4719]. The diurnal alternation of temperatures decreases with burial depth in the soil, and the alternation is also reduced by vegetation cover [@RN1380; @RN4910; @RN4952]. Thus, a positive germination response to alternating temperatures could detect the depth of seed burial, and prevent the germination of seeds that are buried too deep for seedlings to survive before emerging from the ground [@RN4898]. Perhaps more importantly in this habitat, alternating temperatures could also indicate that the vegetation cover has been diminished by either sowing or grazing, and therefore mark an appropriate time for seedling regeneration, when the competition by the established plants is lessened. Similar functions can be ascribed to the germination response to light, as physiologically-active light in significant amounts only travels the uppermost milliliters of soil [@RN4909] and the quality of light will be affected by vegetation cover [@RN604]. Therefore, the germination response to alternating temperatures and light indicate conditions that are found immediately after mowing for haymaking, a predictable perturbation that occurs approximately at the same time every year, and which forces meadow plants to establish, grow and reproduce in the time frame determined by this perturbation [@RN4887; @RN4888]. In addition, these traits would also promote the formation of a persistent soil seed bank [@RN4718; @RN4929; @RN1292]. On the other hand, it is worth mentioning that the Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate meadows [@RN5014; @RN5013].

>>>>

The mesic meadows of Europe are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698]. These plant communities developed in clearings of a temperate woodland landscape as a consequence of anthropogenic intervention [@RN5033; @RN5045]. At a time they were novel plant communities with a composite flora made up of species from different ecological backgrounds, and during historical times their species composition evolved following changes in human activities and management practices [@RN5035; @RN5045]. Mesic meadows are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034]. Nowadays, mesic meadows maintained by agricultural practices are an essential aspect of European landscapes, both real [@RN4996] and imagined [@RN5029].

Traditional classification of mesic grasslands puts emphasis on the separation between meadows used for hay making versus pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893]. Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic meadows in large parts of Europe [@RN4997]. For this reason, the European Habitats Directive (92/43/EEC) has included in its lists of habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.

@@

Mesic meadows are dominated by mesophilous grasses of the family Poaceae, which make up most of the biomass and define the structure of the vegetation. The dispersal unit in Poaceae is generally the floret (**Fig. 1a**), a composite structure made up of the caryopsis or grain (a dry, indehiscent and monospermic fruit) and its surrounding bracts (i.e. modified leaves): the lemma and the palea. Removal of the lemma and palea, as well as puncturing the pericarp, can aid in seed germination [@RN1248]. The dispersal unit is often accompanied by awns or hairs (**Fig. 1b**) that are assumed to aid dispersal, although it is not always the case that they do [@RN1290]. Meadow Poaceae seeds have been reported as having physiological seed dormancy in various degrees [@RN5018; @RN3214; @RN5024], but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273]. Freshly harvested seeds are comparatively more dormant, but dormancy tends to disappear quickly in dry storage [@RN5024; @RN5018]. Germination has been reported to occur at temperatures ranging from 5 to 30 ºC [@RN3273; @RN1229; @RN1261; @RN1230; @RN1292; @RN5024]. Most Poaceae species have also been reported to germinate better in light than in darkness [@RN1271; @RN1248; @RN1230; @RN1221; @RN1172; @RN1292; @RN1280; @RN5024] and even to be inhibited to germinate by darkness [@RN1229]. Nonetheless, species of *Bromus* have been reported as germinating better in darkness [@RN1172] and *Cynosurus cristatus* as being indifferent to light/darkness [@RN1292; @RN1280]. In *Poa trivialis*, germination was promoted by light but not by alternating temperatures [@RN1205], although alternating temperatures did encourage some germination in darkness [@RN1229]. Wild Poaceae ecotypes usually have a germination response to alternating temperatures [@RN1248; @RN1261; @RN1230; @RN1221; @RN1290; @RN1172; @RN1292; @RN1280], although this response is missing in some wild ecotypes and in the domesticated cultivars [@RN1261], and some species such as *Lolium perenne* have been reported as insensitive to temperature alternation [@RN1280; @RN1380].

Next in abundance to Poaceae are the legumes of the family Fabaceae. Fabaceae contribute to the nutritional value of meadow fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040]. The dispersal unit of most Fabaceae is the seed itself (**Fig. 1c**) but in some species dispersal units are more complex, including indehiscent monospermic fruits (e.g. *Onobrychis*, **Fig. 1d**) or indehiscent monospermic fruit fragments, i.e. loments (e.g. *Ornithopus*, **Fig. 1e**). Fabaceae are generally hard-seeded, i.e. they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e. physical seed dormancy) [@RN3214; @RN3273; @RN1215; @RN5016; @RN3073]. In *Medicago*, seeds that have not reached full maturity can germinate before they become impermeable, but the completion of maturation imposes coat impermeability, and thereafter the seed must be scarified to allow water imbibition and germination [@RN652]. Buried Fabaceae seeds are able to track the seasons, and in some species germination seems to be promoted by cold stratification and alternating temperatures [@RN780]; some of these species have been described as having combinational dormancy (i.e. physical + physiological) [@RN5037]. However, Fabaceae seeds have also been reported to germinate without any previous treatment [@RN881; @RN632; @RN460; @RN372], and to lose dormancy during storage [@RN5037]. As in Poaceae, seeds of Fabaceae have been reported to germinate in high numbers across a range of temperatures from 5 to 25 ºC [@RN652; @RN3273]. Fabaceae seeds have been described as not responsive to light and capable of germinating in darkness [@RN1339; @RN3273].

Poaceae and Fabaceae are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the aesthetic value of the meadows [@RN5038; @RN5039]. They also add nutritional scope, being richer than Poaceae and Fabaceae in specific elements [@RN5028; @RN5040]. In many of these families (e.g. Asteraceae, Cyperaceae, Dipsacaceae, Lamiaceae, Polygonaceae, Ranunculaceae) the dispersal unit is the achene: dry, indehiscent and monospermic fruits. The morphology of these achenes is varied: cypselae with a hairy pappus in *Centaurea* (Asteraceae, **Fig. 1f**); beaked in *Ranunculus* (Ranunculaceae, **Fig. 1g**); hardened nutlets in *Prunella* (Lamiaceae, **Fig. 1h**); hairy in *Knautia* (Dipsacaceae, **Fig. 1i**); surrounded by a perigynium which aids in dispersal by water in *Carex* (Cyperaceae, **Fig. 1j**). In the genus *Sanguisorba* (Rosaceae), the dispersal unit is the urn-shaped receptacle containing one to three achenes (**Fig. 1k**). In Apiaceae, it is the mericarp (**Fig. 1l**), an indehiscent monospermic fragment of the fruit. In some other minor families the dispersal unit is the seed itself, such as in Caryophyllaceae (**Fig. 1m**), Plantaginaceae (**Fig. 1n**), Juncaceae or the hemiparasitic species of *Rhinanthus* (Orobanchaceae) (**Fig. 1o**). In Asteraceae, high germination without pretreatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; @RN120; @RN963; @RN1428], *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403]. *Taraxacum officinale* germinates between 5 and 30 ºC [@RN120; RN1428; @RN1276], and has higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428]. In *Stachys officinalis* (Lamiaceae), seeds need either cold stratification, light or alternating temperatures to germinate [@RN416; @RN128]. Underdeveloped embryos that need to grow inside the seed before germination (i.e. morphological dormancy) are widespread in Ranunculaceae and Apiaceae [@RN3214; @RN1112]. *Ranunculus repens* (Ranunculaceae) germinates between 10 and 25 ºC, but the germination percentages were reported to be low [@RN931]; the same species has been reported to respond to alternating temperatures, which can promote its germination even in darkness [@RN3180]. In Polygonaceae, *Rumex acetosa* can germinate immediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not, this difference is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814]. In *Heracleum sphondylium* (Apiaceae), growth of the embryo only occurs below 10 ºC, in moist conditions [@RN1112]. *Sanguisorba minor* (Rosaceae) increases its germination after abrasion of the seeds with bleach [@RN194; @RN226], although germination without pre-treatment has also been reported [@RN298; @RN226]. Seeds of the hemiparasitic species *Rhinanthus angustifolius* and *Rhinanthus minor* (Orobanchaceae) require relatively long periods of cold stratification to germinate [@RN4014], and are able to germinate in the dark [@RN2236].

Although a wealth of studies has accumulated, a synthesis of seed ecology in European mesic meadows is missing. In this article, we review for the first time this topic, combining newly-generated data on seed morphology and germination with records from existing databases [@RN5036; @RN4968; @RN2987; @RN2319]. The resulting dataset contains `r MSrecords` germination records of `r MSspeciesN` plant species from `r MScountries` European countries. To test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light we apply Bayesian meta-analysis [@RN5000]. Further, using well-preserved meadows of the Iberian Peninsula as a study system, we analyze the covariation between seed traits and species environmental preferences, and compare the germination ecology of mesic meadows with that of other herbaceous plant communities from the same area.

@@

Study meadows had been maintained by traditional agricultural practices: mowing for hay-making once or twice per year plus light grazing. Each vegetation sampling plot was placed in a square area (25-100 m^2^ area) situated in the central part of a meadow, avoiding the margins. Vegetation sampling took place in 2016-2017, at the peak of plant development, just before mowing. All vascular plant species in the plots were recorded and given a cover value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%). All plant names were assigned following the nomenclature of Euro+Med [-@RN5044], which is used throughout this article. As expected, the vegetation of the sampled meadows was dominated by Poaceae and Fabaceae: these two families represented 47% and 17%, respectively, of the total plant cover recorded in all the plots. Other 20 families were recorded, the largest of which was Asteraceae; each of these other families represented less than 10% of the total cover.

Using all the relevés, we calculated the cumulative cover of each species in the entire area. To perform the calculation, first we standardized the cover values of the plots by dividing the cover of each species in each plot by the total plant cover in that plot. Then, for each species, we calculated its total cover in the dataset, by summing its standardized cover values from all the plots. Finally, we rescaled the values of all species to a 1-100 scale to obtain the cumulative cover values. We removed 208 species with cumulative cover values below 2%, considering them to be rare species that might have been recorded by chance and may not represent the core mesic meadows flora. We used the remaining 116 species as the core list to retrieve seed germination data for this article. The majority of these species were hemicryptophytes (78%), with some therophytes (16%) and a few chamaephytes and geophytes (3% of each). The family with more species in the list was Poaceae (22%), with another 17% belonging to Fabaceae, 15% to Asteraceae, 8% to Apiaceae, and the rest of the families representing less than 5% each.

We also used the relevés as a basis to characterize the preferences of the selected species for three ecological drivers of mesic meadow diversity: cold, summer drought and soil reaction [@RN2980]. For cold and drought, we used the coordinates of the plots to retrieve from CHELSA [@RN4766] the bioclimatic variables bio06 (minimum temperature of the coldest month) and bio14 (precipitation of the driest month). For soil reaction, we took from each plot five soil samples from between 0 and 20 cm depth with a Dutch auger and combined them to make a bulk soil horizon, which we subsequently air-dried, crumbled, finely crushed and sieved with a 2 mm screen, to finally measure the pH in H~2~O with a glass electrode in a suspension of soil and water (1:2.5). For each of these three environmental variables (bio06, bio14, pH) measured at the plot level we calculated the species niche centroids (SNCs). The SNC for any given species and variable is the mean of the environmental variable in all the plots were the species occurs, weighted by species cover in each plot [@RN4709]. The list of core species, with their cumulative covers and SNCs, is available at GitHub (see Data Availability Statement).

@@

We germinated the collected seeds using three germination treatments to determine the germination response to temperatures that are representative of the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as the summer temperature, and 30/20 ºC as sun-heated soil, e.g. soil exposed to sun after hay cutting. Additionally, we compared, at each of these temperature regimes, the germination of fresh seeds versus seeds subjected to a dormancy-breaking treatment. In the case of Fabaceae and other families that might present physical dormancy [@RN3214], the treatment consisted in scarification by chipping the seed coat with a scalpel. For the rest of the families, we used gibberellic acid GA~3~ (0.0645 mM) in darkness during 24h, as a treatment to remove potential physiological seed dormancy [@RN4976]. For each species and treatment, we sowed four Petri dishes with 25 seeds each. The germination substrate was 1% distilled water - agar. We sealed dishes with Parafilm to prevent desiccation. Trials took place in a germination chamber (KBW 400, Binder GmbH, Tuttlingen, Germany) with a 12/12-hour photoperiod (the light period corresponding to the higher temperature). Experiments lasted for four weeks, with germination scoring once per week. The germination criterion was 2 mm radicle emergence. After four weeks, we cut the seeds that failed to germinate and examined them under a magnifying glass. We classified them as normal when the embryo was visible and firm, empty when they lacked an embryo, and contaminated when they were moldy. We only considered normal seeds to calculate germination proportions and conduct subsequent analyses.

@@

Poaceae had lower values of seed mass and higher values of seed number, while Fabaceae had heavier but fewer seeds (**Fig. 2**). The other families covered the range of values showed by Poaceae and Fabaceae, but their median values were high for both traits: their median seed mass was close to that of the Fabaceae, while their median seed number was higher than that of the Poaceae (**Fig. 2**). Seed shape also showed a divergence between Poaceae and Fabaceae, with seeds being elongated in the former and rounder in the latter (**Fig. 2**). The other families covered the full range of variation, with both elongated and round seeds (**Fig. 2**).

@@

Some differences became apparent when dividing the dataset in the three floristic groups that compose mesic meadow vegetation. In Poaceae (**Fig. 3**), no effect was found for stratification, GA~3~ or temperature. Stratification had a negative effect on germination. The major drivers of Poaceae germination appeared to be alternating temperatures and light, with both having a positive effect. In Fabaceae (**Fig. 3**), the largest positive effect on germination was produced by scarification, with no effect of stratification, GA~3~ or light. Temperature had a negative effect, and in fact Fabaceae species had higher germination proportion at temperatures under 20 ºC. Fabaceae seed germination also responded negatively to alternating temperatures. Finally, in the remaining families (**Fig. 3**), the main cues having a positive effect on germination were stratification, GA~3~, alternating temperatures and light. These species did not respond to scarification nor to average temperature.

@@

PCA indicated a clear separation between environmental preferences and seed traits, with each set of variables contributing to different axes (**Fig. 4**). The first PCA axis explained `r round(pca$eig[1, 2], 0)`% of the variation and was related to environmental preferences. The variables with the largest contribution to this first axis were soil reaction (pH), winter cold (bio06) and summer drought (bio14). This horizontal axis separated (i, left) species with preferences for sites with warm winter temperatures from (ii, right) species with preferences for sites with high summer rainfall and alkaline soils. Axis 2 explained `r round(pca$eig[2, 2], 0)`% of the variability, mostly related to seed traits. The main contributing variables were seed mass, seed number and the germination response to scarification. This axis separated (iii, bottom) Poaceae species that produced many seeds with a positive germination response to alternating temperatures from (iv, top) Fabaceae species that produced heavier seeds with a positive response to scarification.

@@

Ordination resulted in a separation of mesic meadow species from species belonging to natural herbaceous communities of the same region (**Fig. 5**). Axis 1 explained 78% of the variance and was positively associated to high germination at all three temperature treatments. This horizontal axis separated (i, left) species with low germination across treatments from (ii, right) species with high germinability across treatments. The second axis explained 17% of the variation and separated (iii, bottom) species that responded more to the cool temperature germination treatment from (iv, top) species that responded more positively to the warm germination treatment. Meadow species tended to be situated in the right sector of the ordination, indicating high germinability; and at the center of the vertical axis, indicating a neutral response to temperature. On the other hand, species from the other communities were positioned at the left of the horizontal axis, indicating a lower germinability; and were more separated along the vertical axis, indicating a preference for either warmer (bogs and fens) or cooler (alpine grasslands, coastal communities) germination treatments.

@@

Our meta-analysis of germination records showed that, overall, the regeneration of mesic meadows is characterized by (1) high germinability when compared to other herbaceous plant communities; (2) low correspondence between seed trait variability and the natural environmental drivers of mesic meadow diversity; and (3) a strong phylogenetic separation between the two major families, Poaceae and Fabaceae.

We found that alternating temperatures and light are the two most influential factors triggering germination, especially in species from the family Poaceae and the remaining ones, with the exception of Fabaceae. Alternating temperatures and light are usually considered to be micro-environmental cues that indicate, at a fine scale, the existence of safe sites for regeneration [@RN4719]. The diurnal alternation of temperatures decreases with burial depth in the soil, and the alternation is also reduced by vegetation cover [@RN1380; @RN4910; @RN4952]. Thus, a positive germination response to alternating temperatures could detect the depth of seed burial, and prevent the germination of seeds that are buried too deep for seedlings to survive before emerging from the ground [@RN4898]. Perhaps more importantly in this habitat, alternating temperatures could also indicate that the vegetation cover has been diminished by either sowing or grazing, and therefore mark an appropriate time for seedling regeneration, when the competition by the established plants is lessened. Similar functions can be ascribed to the germination response to light, as physiologically-active light in significant amounts only travels the uppermost milliliters of soil [@RN4909] and the quality of light will be affected by vegetation cover [@RN604]. Therefore, the germination response to alternating temperatures and light indicate conditions that are found immediately after mowing for haymaking, a predictable perturbation that occurs approximately at the same time every year, and which forces meadow plants to establish, grow and reproduce in the time frame determined by this perturbation [@RN4887; @RN4888]. In addition, these traits would also promote the formation of a persistent soil seed bank [@RN4718; @RN4929; @RN1292]. On the other hand, it is worth mentioning that Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate meadows [@RN5014; @RN5013].
